---
title: "Yeager_Lab3"
author: "Joel Yeager"
date: "2026-02-26"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
```

## Lab 3: Conjugate Priors

For each scenario below:

-   a\. Identify appropriate conjugate distributions

-   b\. Plot both the posterior and the prior distributions for the modeled quantity and interpret in the same figure.

-   c\. Using the posterior distribution, determine the 95% credible interval for the parameter in each scenario.

-   d\. Comment on the informative nature of the prior in terms of its relationship to the posterior.

***Scenarios***

1.  Climate Referendum Voting

    You are conducting a political opinion poll and have data from 100 voters at random. Each voter whether they will vote “yes” in an upcoming referendum on funding for local climate adaptation. Little is known about how people might vote. Estimate the probability that the referendum passes. The poll results showed that 38 people voted yes. Answer questions a-d. In addition, estimate the probability that the referendum passes as Pr(ϕ≥.5).

```{r}
y <- 38
n <- 100
aPrior <- 1
bPrior <- 1


(aPost <- aPrior + y)
(bPost <- bPrior + n - y)
post <- beta(39,63)

# A) Prior: uniform beta; Liklihood: binomial; Posterior: Beta-binomial

# B) Plot posterior and prior:

```

2.  Lobster Catch!

    A group of fishermen go lobster fishing for an hour. The fishermen catch 15, 8, 6, 11, and 4 lobsters. Estimate average catch, given that, prior to this fishing trip, we believed fishermen catch an average of 3 lobsters per hour (μ=3) an hour with a standard deviation of 1.5 lobsters per hour (σ=1.5).

```{r}

```

3.  Solstice temperature

    You have collected temperature data on the December solstice for the last 50 years in Maryland. The known standard deviation for these measurements is 15 (σ=15). In a previous study researchers found mean temperature to be 30 (μprior=30) with a standard deviation of 12 (σprior=12). The solstice temperature measurements are stored in the data SolsticeTemp.csv.

```{r}
temps <- read_csv('SolsticeTemp.csv')
```
